---
title: "MGS for Sharon"
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 80    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
getwd()
library(usethis)
library(renv)
library(RMark)
library(tictoc)
library(dplyr)
library(rmarkdown)

cat("
Ran at: ");Sys.time()
```

Download data

```{r}
tic()
mgs.cjsdata <- read.table(file = "mgs.cgs.site1.txt", header = TRUE, colClasses = c("character","factor","factor"), col.names = c("ch", "captage", "sex"))

#sort the data so we know how it is arranged
arrange(mgs.cjsdata, mgs.cjsdata$sex, mgs.cjsdata$captage)

# this appears to just let you view a portion of it
summary(mgs.cjsdata)

cat("
Ran at: ");Sys.time()
toc()
```

subset the data so we can do a CJS analysis instead of Multistrata

```{r}
tic()

#replace a, b and c with 1s
mgs.cjsdata$ch <- gsub("2","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("3","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("4","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("5","1", mgs.cjsdata$ch)

cat("
Ran at: ");Sys.time()
toc()
```

Process the data this step sets up the data for the appropriate model
type in MARK Laake uses the term dp in this step. I am using pd for
process(ed) data

```{r}
tic()
#code for multistrata analysis
#mgs.pd <- process.data(mgs.data1, model = "Multistrata", groups = c("captage", "sex"))

mgs.pd <- process.data(mgs.cjsdata, model = "CJS", groups = c("sex", "captage"))

str(mgs.pd)

cat("
Ran at: ");Sys.time()
toc()
```

Create the files needed for analysis in MARK This is where design
matrices are created Laake uses the term ddl for design data list in
this step

For CJS data - for 500 rows of data - observations all converted to 1s
This took a 98 seconds to run

For MS data - Additional steps will be needed to set up the appropriate
dm for age and movement parameters With this dataset, this takes a long
time to run (I stopped it after an hour)

```{r}
tic()
mgs.ddl <- make.design.data(mgs.pd)

summary(mgs.ddl)

cat("
Ran at: ");Sys.time()
toc()
```

Run the default model in MARK this will be model 1 output m1o

```{r}
tic()

m1o <- mark(mgs.pd, mgs.ddl)

str(m1o)

cat("
Ran at: ");Sys.time()
toc()
```

Real estimates from model 1
```{r}
m1o$results$real

cat("
Ran at: ");Sys.time()
```

Setup design matrices start by reading days since start

```{r}
tic()
ti1 <- (read.csv("M:/Cloud_storage/csg-google-drive/Sharon_MGS/cjs.time.interVals.csv", header = FALSE))
ti.yr1 <- ti1/365
ti <- as.numeric(ti1[1:328,])
ti.yr <- as.numeric(ti.yr1[1:328,])
summary(ti.yr)

#cleanup
rm(ti1)
rm(ti.yr1)

cat("
Ran at: ");Sys.time()
toc()
```

add correct time interval - in years - to the process data

```{r}
tic()
mgs.pd$time.intervals <- ti.yr

#rerun model 1 - output = FALSE just doesn't print the output
m1o <- mark(mgs.pd, mgs.ddl, output = FALSE)

m1o$results$real

cat("
Ran at: ");Sys.time()
toc()
```

Set initial ages
and try to set up design matrices for age-specific survival 
Mark uses convention of capital letters to designate continuous variables and lower case for factors (e.g., Age and age)
```{r}
tic()

#top of page 46 RMark Workshop Notes

# set number of age variable to 1 (first group variable in list of groups), set initial ages (by captage group Alphabetically)

mgs2.pd <- process.data(mgs.cjsdata, model = "CJS", groups = c("captage", "sex"), age.var = 1, initial.ages = c(1,0), time.intervals = ti.yr)

mgs2.ddl <- make.design.data(mgs2.pd)

#change age to a factor of 1 or 0 for adult or juv

mgs2.ddl$Phi$age = factor(ifelse(mgs2.ddl$Phi$Age<1,0,1 ))
mgs2.ddl$p$age = factor(ifelse(mgs2.ddl$p$Age<1,0,1))

cat("
Ran at: ");Sys.time()
toc()
```




Rerun model with age-specific survival+

```{r}
tic()

Phi.age=list(formula=~age)
p.age = list(formula=~age)

m2o <- mark(mgs2.pd, mgs2.ddl, model.parameters = list(Phi=Phi.age, p = p.age), output = FALSE)



m2o$results$real



cat("
Ran at: ");Sys.time()
toc()
```






```{r}
tic()
#define some formulas for models
Phi.age <- list(formula=~age)
p. <- list(formula=~1)

m3o <- mark(mgs2.pd, mgs2.ddl, model.parameters = list(Phi=Phi.age, p = p.), output = FALSE)

m3o$results$real

cat("
Ran at: ");Sys.time()
toc()
```








Now try to run a set of different models and look at the results
Page 70 of Rmark workshop notes

This list of 9 models took 3 minutes to run

mgs.mso is mgs model set output

```{r}
tic()

mgs.analysis = function(){
  #formulas for Phi
Phi.    = list(formula=~1     )
Phi.sex = list(formula=~sex   )
Phi.age = list(formula=~age   )
  #formulas for p
p.    = list(formula=~1     )
p.sex = list(formula=~sex   )
p.age = list(formula=~age     )
  #create model list (ml) of all combinations of models
mgs.ms=create.model.list("CJS")
  # trying to speed things up using parallel processing
#results = mark.wrapper.parallel(mgs.ms, data = mgs.pd, ddl = mgs.ddl, output = FALSE, silent = TRUE, parallel = TRUE, cpus = 16, threads = 1)
results = mark.wrapper(mgs.ms, data = mgs2.pd, ddl = mgs2.ddl, output = FALSE, silent = TRUE )
return(results)
}

mgs.mso = mgs.analysis()

mgs.mso

cat("
Ran at: ");Sys.time()
toc()
```


Run analysis as robust design for site 2

Read data files 
```{r}
tic()
mgs.rd.data.site2a <- read.table(file = "mgs.rd.site2.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor"), col.names = c("ch", "captage", "sex", "id") )

mgs.rd.data.site2 <- subset(mgs.rd.data.site2a, mgs.rd.data.site2a$id != "Dummy" )

#replace strata with 1s
mgs.rd.data.site2$ch <- gsub("2","1", mgs.rd.data.site2$ch)
mgs.rd.data.site2$ch <- gsub("3","1", mgs.rd.data.site2$ch)
mgs.rd.data.site2$ch <- gsub("4","1", mgs.rd.data.site2$ch)

ti.site2a <- (read.csv("mgs.rd.time.intervals.site2.csv", header = FALSE))
ti.site2 <- as.numeric(ti.site2a[1:709,])

mgs.time.intervals.rd.yr <- ti.site2/365

write.csv(mgs.time.intervals.rd.yr,"rd.yr.csv")

#cleanup
rm(mgs.rd.data.site2a)
rm(ti.site2a)


cat("
Ran at: ");Sys.time()
toc()
```


Set up analysis

Be careful with the groups that you define in pd.  Too many groups will cause make.design.data take a long time to process.

```{r}
tic()
mgs.rd.pd <- process.data(data = mgs.rd.data.site2, model = "Robust", groups = c("captage", "sex"), time.intervals = mgs.time.intervals.rd.yr, age.var = 1, initial.ages = c(1,0), age.unit = 1)

mgs.rd.ddl <- make.design.data(mgs.rd.pd)


mgs.rd.ddl$S$age = factor(ifelse(mgs.rd.ddl$S$Age<1,0,1 ))


cat("
Ran at: ");Sys.time()
toc()

```




Run a model set
This set of models took 45 minutes to run when I didn't specify models for f0
Took 3 minutes with only 3 survival models 
took 3 1/2 minutes with 4 survival models
f0 are the estimate of N 

```{r}
tic()

mgs.analysis = function(){
  #formulas for S
S.dot    = list(formula=~1     )
S.sex = list(formula=~sex   )
S.age = list(formula=~age   )
S.agesex = list(formula=~age + sex)
  #formulas for p c gamma` gamma ``
p.dot = list(formula=~1)
c.dot = list(formula=~1)
GammaDoublePrime.dot = list(formula=~1)
GammaPrime.dot = list(formula=~1)
f0.dot = list(formula=~1)
  #create model list (ml) of all combinations of models
mgs.ms=create.model.list("Robust")
  # trying to speed things up using parallel processing
#results = mark.wrapper.parallel(mgs.ms, data = mgs.rd.data.site2, ddl = mgs.rd.ddl, output = FALSE, silent = TRUE, parallel = TRUE, cpus = 16, threads = -1)
results = mark.wrapper(mgs.ms, data = mgs.rd.pd, ddl = mgs.rd.ddl, output = FALSE, silent = TRUE )
return(results)
}

mgs.mso = mgs.analysis()

mgs.mso

cat("
Ran at: ");Sys.time()
toc()
```
Look at results for each model 
```{r}
mgs.mso$S.agesex.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot.f0.dot$results$real
mgs.mso$S.age.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot$results$real
mgs.mso$S.sex.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot$results$real
mgs.mso$S.dot.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot$results$real

```

Run analysis for all 4 sites

Download data for all sites 
time intervals are identical for all sites


```{r}
tic()
mgs.rd.data.site1a <- read.table(file = "mgs.rd.site1.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site2a <- read.table(file = "mgs.rd.site2.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site3a <- read.table(file = "mgs.rd.site3.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site4a <- read.table(file = "mgs.rd.site4.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )

mgs.rd.alla <- rbind(mgs.rd.data.site1a, mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)

mgs.rd.allb <- subset(mgs.rd.alla, mgs.rd.alla$id != "Dummy" )
mgs.rd.allc <- subset(mgs.rd.allb, mgs.rd.allb$captage != "" )

#replace strata with 1s
mgs.rd.allc$ch <- gsub("2","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("3","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("4","1", mgs.rd.allc$ch)

mgs.rd.allc$captage <- as.factor(ifelse(mgs.rd.allc$captage1=="ADU",1,0))

#only include certain columns of data
mgs.rd.all <- mgs.rd.allc[c(1,3,5,6)]

#possibly something wrong here 12-16-22
ti.site2 <- as.numeric(read.csv("mgs.rd.time.intervals.site2.csv", header = FALSE))
mgs.time.intervals.rd.yr <- ti.site2/365

#cleanup
rm(mgs.rd.alla, mgs.rd.allb, mgs.rd.allc, ti.site2, mgs.rd.data.site1a, mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)


cat("
Ran at: ");Sys.time()
toc()

```


```{r}
tic()
mgs.rd.pd <- process.data(data = mgs.rd.all, model = "Robust", groups = c("sex","captage", "site"), time.intervals = mgs.time.intervals.rd.yr, age.var = 2, initial.ages = c(1,0), age.unit = 1)

mgs.rd.ddl <- make.design.data(mgs.rd.pd)


mgs.rd.ddl$S$age = factor(ifelse(mgs.rd.ddl$S$Age<1,0,1 ))


cat("
Ran at: ");Sys.time()
toc()

```

Run model set 
This took 75 minutes
```{r}
tic()

mgs.analysis = function(){
  #formulas for S
S.dot    = list(formula=~1     )
S.sex = list(formula=~sex   )
S.age = list(formula=~age   )
S.site = list(formula=~site)
S.agesex = list(formula=~age + sex)
S.agesite = list(formula=~age + site)
S.sexsite = list(formula=~sex + site)
S.ass = list(formula=~age + sex + site)
  #formulas for p c gamma` gamma ``
p.dot = list(formula=~1)
c.dot = list(formula=~1)
GammaDoublePrime.dot = list(formula=~1)
GammaPrime.dot = list(formula=~1)
f0.dot = list(formula=~1)
  #create model list (ml) of all combinations of models
mgs.ms=create.model.list("Robust")
  # trying to speed things up using parallel processing
#results = mark.wrapper.parallel(mgs.ms, data = mgs.rd.data.site2, ddl = mgs.rd.ddl, output = FALSE, silent = TRUE, parallel = TRUE, cpus = 16, threads = -1)
results = mark.wrapper(mgs.ms, data = mgs.rd.pd, ddl = mgs.rd.ddl, output = FALSE, silent = TRUE )
return(results)
}

mgs.mso = mgs.analysis()

mgs.mso

cat("
Ran at: ");Sys.time()
toc()
```





```{r}
tic()

write.csv(mgs.mso$S.sexsite.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot.f0.dot$results$real, file = "S.sex site.csv")
write.csv(mgs.mso$S.ass.GammaDoublePrime.dot.GammaPrime.dot.p.dot.c.dot.f0.dot$results$real, file = "S.age sex site.csv")

cat("
Ran at: ");Sys.time()
toc()
```


Clean up the folder - remove unneeded mark files. 

```{r}
cleanup(ask = FALSE)
```




___________________________________________________

Include data to constrain p and c by site and date

Here to next section defined by red lines is a complete analysis
_______________________________________________________
```{r}
tic()
#input data for each site
mgs.rd.data.site1a <- read.table(file = "mgs.rd.site1.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site2a <- read.table(file = "mgs.rd.site2.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site3a <- read.table(file = "mgs.rd.site3.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site4a <- read.table(file = "mgs.rd.site4.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )

mgs.rd.alla <- rbind(mgs.rd.data.site1a, mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)

mgs.rd.allb <- subset(mgs.rd.alla, mgs.rd.alla$id != "Dummy" )
mgs.rd.allc <- subset(mgs.rd.allb, mgs.rd.allb$captage != "" )

#replace strata with 1s
mgs.rd.allc$ch <- gsub("2","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("3","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("4","1", mgs.rd.allc$ch)

mgs.rd.allc$captage <- as.factor(ifelse(mgs.rd.allc$captage=="ADU",1,0))

mgs.rd.all <- mgs.rd.allc

#cleanup 

rm(mgs.rd.alla, mgs.rd.allb, mgs.rd.allc, mgs.rd.data.site1a ,mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)

# input the data for each site to include the constrained values for p and c by date since start
# days since start are only for the primary occassions (sessions) 
fixed.pc <- read.csv("fixed.pc.csv", header = TRUE, fill = TRUE)
site1.fixed.p <- fixed.pc[,"site1.fixed.p"]
site2.fixed.p <- fixed.pc[,"site2.fixed.p"]
site3.fixed.p <- fixed.pc[,"site3.fixed.p"]
site4.fixed.p <- fixed.pc[,"site4.fixed.p"]

# set fixed p for 16 groups where site is last - 
# 710 lines * 4 = 2840 for site 1, next 2840 site 2, etc.
fixed.p <- c(rep(site1.fixed.p, 4), rep(site2.fixed.p,4), rep(site3.fixed.p, 4), rep(site4.fixed.p,4))

#set fixed for c similarly except do not include the "s"
a <- fixed.pc[,"site1.fixed.c"]
site1.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site2.fixed.c"]
site2.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site3.fixed.c"]
site3.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site4.fixed.c"]
site4.fixed.c <- as.numeric(subset(a, a!="s"))

fixed.c <- c(rep(site1.fixed.c,4), rep(site2.fixed.c,4), rep(site3.fixed.c,4), rep(site4.fixed.c,4))

# set time intervals 
a <- fixed.pc[,"Time.intervals..years."]
rd.yr <- as.numeric(a[1:709])

#cleanup 
rm(a, site1.fixed.c, site1.fixed.p, site2.fixed.c, site2.fixed.p, site3.fixed.c, site3.fixed.p, site4.fixed.c, site4.fixed.p)


cat("
Ran at: ");Sys.time()
toc()
```





process data and make ddl



```{r}
tic()

mgs.rd.pd <- process.data(data = mgs.rd.all, model = "Robust", groups = c("sex","captage", "site"), time.intervals = rd.yr, age.var = 2, initial.ages = c(1,0), age.unit = 1)

mgs.rd.ddl <- make.design.data(mgs.rd.pd)


mgs.rd.ddl$S$age = factor(ifelse(mgs.rd.ddl$S$Age<1,0,1 ))




cat("
Ran at: ");Sys.time()
toc()
```



Clean up the folder - remove unneeded mark files. 

```{r}
cleanup(ask = FALSE)
```




Run a model with constraints



```{r}
tic()
S.ass = list(formula=~age + sex + site)
g.site = list(formula=~site)
p.fix = list(formula=~1, fixed=fixed.p)
c.fix = list(formula=~1, fixed=fixed.c)
f0.site = list(formula=~site)
mrd1 <- mark(mgs.rd.pd, mgs.rd.ddl, model.parameters = list(S = S.ass, GammaDoublePrime = g.dot, GammaPrime = g.dot, p=p.fix, c=c.fix, f0=f0.site))


cat("
Ran at: ");Sys.time()
toc()
```



```{r}
tic()

mrd1$results$real



write.csv(mrd1$results$real, file = "Real_fixed_pc.csv")
write.csv(mrd1$results$beta, file = "Beta_fixed_pc.csv")



cat("
Ran at: ");Sys.time()
toc()
```

______________________________________________________________

Modify code to add time-specific and site-specific covariates
________________________________________________________________


Clean up the folder - remove unneeded mark files. 

```{r}
tic()

cleanup(ask = FALSE)

cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min")
```





First - copy code from last successful run to add to it. 

Same as lines 501-565

```{r}
tic()
#input data for each site
mgs.rd.data.site1a <- read.table(file = "mgs.rd.site1.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site2a <- read.table(file = "mgs.rd.site2.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site3a <- read.table(file = "mgs.rd.site3.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )
mgs.rd.data.site4a <- read.table(file = "mgs.rd.site4.txt", header = TRUE, colClasses = c("character", "factor", "factor", "factor", "factor"), col.names = c("ch", "captage1", "sex", "id", "site") )

mgs.rd.alla <- rbind(mgs.rd.data.site1a, mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)

mgs.rd.allb <- subset(mgs.rd.alla, mgs.rd.alla$id != "Dummy" )
mgs.rd.allc <- subset(mgs.rd.allb, mgs.rd.allb$captage != "" )

#replace strata with 1s
mgs.rd.allc$ch <- gsub("2","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("3","1", mgs.rd.allc$ch)
mgs.rd.allc$ch <- gsub("4","1", mgs.rd.allc$ch)

mgs.rd.allc$captage <- as.factor(ifelse(mgs.rd.allc$captage=="ADU",1,0))

mgs.rd.all <- mgs.rd.allc

#cleanup 

rm(mgs.rd.alla, mgs.rd.allb, mgs.rd.allc, mgs.rd.data.site1a ,mgs.rd.data.site2a, mgs.rd.data.site3a, mgs.rd.data.site4a)

# input the data for each site to include the constrained values for p and c by date since start
# days since start are only for the primary occassions (sessions) 
fixed.pc <- read.csv("fixed.pc.csv", header = TRUE, fill = TRUE)
site1.fixed.p <- fixed.pc[,"site1.fixed.p"]
site2.fixed.p <- fixed.pc[,"site2.fixed.p"]
site3.fixed.p <- fixed.pc[,"site3.fixed.p"]
site4.fixed.p <- fixed.pc[,"site4.fixed.p"]

# set fixed p for 16 groups where site is last - 
# 710 lines * 4 = 2840 for site 1, next 2840 site 2, etc.
fixed.p <- c(rep(site1.fixed.p, 4), rep(site2.fixed.p,4), rep(site3.fixed.p, 4), rep(site4.fixed.p,4))

#set fixed for c similarly except do not include the "s"
a <- fixed.pc[,"site1.fixed.c"]
site1.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site2.fixed.c"]
site2.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site3.fixed.c"]
site3.fixed.c <- as.numeric(subset(a, a!="s"))

a <- fixed.pc[,"site4.fixed.c"]
site4.fixed.c <- as.numeric(subset(a, a!="s"))

fixed.c <- c(rep(site1.fixed.c,4), rep(site2.fixed.c,4), rep(site3.fixed.c,4), rep(site4.fixed.c,4))

# set time intervals 
a <- fixed.pc[,"Time.intervals..years."]
rd.yr <- as.numeric(a[1:709])

#cleanup 
rm(a, site1.fixed.c, site1.fixed.p, site2.fixed.c, site2.fixed.p, site3.fixed.c, site3.fixed.p, site4.fixed.c, site4.fixed.p)


cat("
Ran at: ");Sys.time()
e = toc() 
cat((e$toc-e$tic)/60, "min")
```
Add site-specific covariates as individual covariates
elevation
```{r}
tic()

mgs.rd.all$elevation <- as.numeric(
  ifelse(mgs.rd.all$site == 1, 1024, 
         ifelse(mgs.rd.all$site == 2, 1094, 
                ifelse(mgs.rd.all$site == 3, 1475,
                       ifelse(mgs.rd.all$site ==4, 1513, 0))))
)



cat("
Ran at: ");Sys.time()
e = toc() 
cat((e$toc-e$tic)/60, "min")
```

set up analysis using elevation as covariate for survival


```{r}
tic()


mgs.rd.pd <- process.data(data = mgs.rd.all, model = "Robust", groups = c("sex","captage", "site"), time.intervals = rd.yr, age.var = 2, initial.ages = c(1,0), age.unit = 1)

mgs.rd.ddl <- make.design.data(mgs.rd.pd)


mgs.rd.ddl$S$age = factor(ifelse(mgs.rd.ddl$S$Age<1,0,1 ))



cat("
Ran at: ");Sys.time()
e = toc() 
cat((e$toc-e$tic)/60, "min")
```



Run the model 
says 16 minutes but that is not true

```{r}
tic()

S.ae = list(formula=~age + elevation)
g.site = list(formula=~site)
p.fix = list(formula=~1, fixed=fixed.p)
c.fix = list(formula=~1, fixed=fixed.c)
f0.site = list(formula=~site)
mrd1 <- mark(mgs.rd.pd, mgs.rd.ddl, model.parameters = list(S = S.ae, GammaDoublePrime = g.site, GammaPrime = g.site, p=p.fix, c=c.fix, f0=f0.site))




cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min")
```


Show results

```{r}
start <- Sys.time()
tic()

real <- mrd1$results$real
real

write.csv(real, "real.csv")


cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min
")

end <- Sys.time()
cat("Elapsed time: "); end-start

```



```{r}
start <- Sys.time()
tic()

cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min
")

end <- Sys.time()
cat("Elapsed time: "); end-start

```



```{r}
start <- Sys.time()
tic()

cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min
")

end <- Sys.time()
cat("Elapsed time: "); end-start

```








Copy and paste r chunk to include timing information


```{r}
start <- Sys.time()
tic()

cat("
Ran at: ");Sys.time()
e = toc()
cat((e$toc-e$tic)/60, "min
")

end <- Sys.time()
cat("Elapsed time: "); end-start

```







HAVEN'T SUCCESSFULLY RUN CODE BELOW



Export data to process directly in MARK


```{r}
tic()

m2=mark(mgs.pd, mgs.ddl, model.parameters = list(Phi=p., p = p.), output = FALSE)

export.MARK(mgs.pd, "mgs.cjsdata", m2, replace = TRUE )


cat("
Ran at: ");Sys.time()
toc()
```


