---
title: "MGS for Sharon"
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 80    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
getwd()

library(RMark)
library(tictoc)
library(dplyr)
library(rmarkdown)

cat("
Ran at: ");Sys.time()
```

Download data

```{r}
tic()
mgs.cjsdata <- read.table(file = "mgs.cgs.site1.txt", header = TRUE, colClasses = c("character","factor","factor"), col.names = c("ch", "captage", "sex"))

#sort the data so we know how it is arraged
arrange(mgs.cjsdata, mgs.cjsdata$sex, mgs.cjsdata$captage)

# this appears to just let you view a portion of it
summary(mgs.cjsdata)

cat("
Ran at: ");Sys.time()
toc()
```

subset the data so we can do a CJS analysis instead of Multistrata

```{r}
tic()

#replace a, b and c with 1s
mgs.cjsdata$ch <- gsub("2","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("3","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("4","1", mgs.cjsdata$ch)
mgs.cjsdata$ch <- gsub("5","1", mgs.cjsdata$ch)

cat("
Ran at: ");Sys.time()
toc()
```

Process the data this step sets up the data for the appropriate model
type in MARK Laake uses the term dp in this step. I am using pd for
process(ed) data

```{r}
tic()
#code for multistrata analysis
#mgs.pd <- process.data(mgs.data1, model = "Multistrata", groups = c("captage", "sex"))

mgs.pd <- process.data(mgs.cjsdata, model = "CJS", groups = c("sex", "captage"))

str(mgs.pd)

cat("
Ran at: ");Sys.time()
toc()
```

Create the files needed for analysis in MARK This is where design
matrices are created Laake uses the term ddl for design data list in
this step

For CJS data - for 500 rows of data - observations all converted to 1s
This took a 98 seconds to run

For MS data - Additional steps will be needed to set up the appropriate
dm for age and movement parameters With this dataset, this takes a long
time to run (I stopped it after an hour)

```{r}
tic()
mgs.ddl <- make.design.data(mgs.pd)

summary(mgs.ddl)

cat("
Ran at: ");Sys.time()
toc()
```

Run the default model in MARK this will be model 1 output m1o

```{r}
tic()

m1o <- mark(mgs.pd, mgs.ddl)

str(m1o)

cat("
Ran at: ");Sys.time()
toc()
```

Real estimates from model 1
```{r}
m1o$results$real

cat("
Ran at: ");Sys.time()
```

Setup design matrices start by reading days since start

```{r}
tic()
ti1 <- (read.csv("M:/CSG Google Drive/Sharon MGS/cjs.time.interVals.csv", header = FALSE))
ti.yr1 <- ti1/365
ti <- as.numeric(ti1[1:328,])
ti.yr <- as.numeric(ti.yr1[1:328,])
summary(ti.yr)

#cleanup
rm(ti1)
rm(ti.yr1)

cat("
Ran at: ");Sys.time()
toc()
```

add correct time interval - in years - to the process data

```{r}
tic()
mgs.pd$time.intervals <- ti.yr

#rerun model 1 - output = FALSE just doesn't print the output
m1o <- mark(mgs.pd, mgs.ddl, output = FALSE)

m1o$results$real

cat("
Ran at: ");Sys.time()
toc()
```

Set initial ages
and try to set up design matrices for age-specific survival 
Mark uses convention of capital letters to designate continuous variables and lower case for factors (e.g., Age and age)
```{r}
tic()

#top of page 46 RMark Workshop Notes
# set number of age variables to 2, set initial ages (by captage group Alphabetically)
mgs.pd <- process.data(mgs.cjsdata, model = "CJS", groups = c("captage", "sex"), age.var = 2, initial.ages = c(1,0), time.intervals = ti.yr)

str(mgs.pd)

cat("
Ran at: ");Sys.time()
toc()
```




THIS DIDN'T WORK 

```{r}
tic()
#defining age bins so the bins are 0 to 1, 1 to 50 and 50 to? Really the last bin doesn't exist and 50 is > max age
# right = TRUE meand the bins are closed on the right
mgs.ddl <- make.design.data(mgs.pd, parameters = list(Phi=list(age.bins=c(0,1,50))), right = TRUE)
levels(mgs.ddl$Phi$age) = c("1plus","0")
summary(mgs.ddl$Phi$age)

cat("
Ran at: ");Sys.time()
toc()
```


```{r}
tic()
#define some formulas for models
Phi.age <- list(formula=~age)
p. <- list(formula=~1)

m1o <- mark(mgs.pd, mgs.ddl, model.parameters = list(Phi=Phi.age, p = p.), output = FALSE)

m1o$results$real

cat("
Ran at: ");Sys.time()
toc()
```
Export data to process directly in MARK


```{r}
tic()

m2=mark(mgs.pd, mgs.ddl, model.parameters = list(Phi=p., p = p.), output = FALSE)

export.MARK(mgs.pd, "mgs.cjsdata", m2, replace = TRUE )


cat("
Ran at: ");Sys.time()
toc()
```




Now try to run a set of different models and look at the results
Page 70 of Rmark workshop notes
THIS CAUSED ME PROBLEMS - ENDED UP RUNNING FOR AN HOUR AND HALF AND USING 95% OF MY 96GB memory

```{r}
tic()

mgs.analysis = function(){
  #formulas for Phi
Phi.    = list(formula=~1     )
Phi.sex = list(formula=~sex   )
Phi.age = list(formula=~age   )
  #formulas for p
p.    = list(formula=~1     )
p.sex = list(formula=~sex   )
p.age = list(formula=~age     )
  #create model list (ml) of all combinations of models
mgs.ms=create.model.list("CJS")
  # trying to speed things up using parallel processing
#results = mark.wrapper.parallel(mgs.ms, data = mgs.pd, ddl = mgs.ddl, output = FALSE, silent = TRUE, parallel = TRUE, cpus = 16, threads = 1)
results = mark.wrapper(mgs.ms, data = mgs.pd, ddl = mgs.ddl, output = FALSE, silent = TRUE )
return(results)
}

mgs.mso = mgs.analysis()

mgs.mso

cat("
Ran at: ");Sys.time()
toc()
```




Copy and paste r chunk to include timing information


```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```

